### 2022-09-28 src/tex/intro.tex
Add comments and update abs/into/methods -- add o3b gaussian chieff to plot

\section{Introduction} \label{sec:intro}

Observations of gravitational waves (GWs) from compact binary coalescences (CBCs) are becoming a regular occurrence, 
recently producing catalogs of 90 such detections \citep{GWTC1,gwtc2,GWTC3}. With this growing catalog size, we are steadily approaching a more complete  
understanding of the population properties of CBCs, just 7 years after the first observation \citep{o1o2_pop,o3a_pop,o3b_astro_dist}. 
In the coming fourth observing run of the LIGO-Virgo-KAGRA (LVK) collaboration detectors \citep{aLIGO,aVIRGO}, with the increased sensitivities, 
there could be a detection as often as every day \citep{LVK_prospects}. This large influx of incoming data will drastically increase what we can 
learn by combining all observations in order to study the compact object population. The history of these dense objects is imprinted on the properties 
of the population  of CBCs, allowing us to learn about their formation and evolution by the last few seconds of gravitational radiation emitted \citep{Zevin_2017}. 
The population distributions of merging compact objects inferred from gravitational-wave observations have been shown to be particularly useful in order to
measure cosmological parameters \citep{Farr_2019HUB,gwtc3_cosmo}, constrain modified gravitational wave propagation \citep{ModGWProp,ModGWProp2}, 
constrain a running Planck mass \citep{Lagos_runningPlanckMass}, look for evidence of ultralight bosons through superradiance \citep{Ng_Boson2021,GWTC2_superradiance_Ng}, 
constrain stellar nuclear reaction rates \citep{Farmer_2019,Farmer_2020}, look for primordial black holes \citep{Ng_2021,KenNgPBH2022}, 
and constrain the neutron star equation of state \citep{Golomb_EOS}. Through a better understanding of the mass, spin, and redshift distributions of 
compact binaries that will come with the increased catalog size, one can probe a wide range of different physical phenomena with even greater fidelity.

Studies using flexible methods to analyze the binary black hole (BBH) mass distribution with GWTC-2 found signs of structure 
beyond the first reported feature found at $m_1\sim35\,\msun$ \citep{Talbot_2018,o3a_pop,Tiwari_2021_b,Edelman_2022ApJ}. The evidence for this structure, 
such as the peak at $m_1\sim10\,\msun$, has only increased after analyzing the following catalog, GWTC-3 \citep{o3b_astro_dist,Tiwari_2022ApJ}. 
This substructure was first confidently found with multiple different flexible or non-parametric models that are data-driven with little 
bias on the functional form. Conversely, the other canonical approach is to use simple parametric models that aim to describe the data
in the simplest way, building in astrophysically motivated priors, followed by sequentially adding complexities only as the data demands it. While in data-scarce, prior dominated
environments, this simple approach is more powerful. However, as we go forward into the data-rich catalog era, flexible models will be necessary to discover new features that may be in
the data, which quickly becomes unwieldy with the sequential approach. Non-parametric or semi-parametric methods are particularly useful in looking for unexpected features in the data, finding where 
our simpler parametric models may be failing to capture meaningful features in the data. Development and robust testing of such methods in the short term will provide
knock on effects as the size of our catalogs increase, and GW astronomy enters its data-dominated era.

The spin distributions of BBHs are particularly interesting since the two main formation scenarios produce distinctly different spin distributions \bruce{cite}. 
Isolated or field formation scenarios predict component spins that are preferentially aligned with the binaries orbital angular momentum, although some small 
misalignment can occur depending on the nature of the supernova kicks as each star collapses to a compact object \bruce{cite}. Dynamical formation scenarios may 
occur in dense environments where many-body interactions between a large number compact objects occur during the binary formation hardening the binary so that it 
will merge within a Hubble time. Thus, the dynamical channel produces an isotropic spin tilt distribution \bruce{cite}. BBH spins have also been of controversial 
interest recently, with different parametric approaches to modeling the spin distribution coming to different conclusions. The studies have disagreed on the possible existence 
of a significant zero-spinning subpopulation, along with the presence of significant spin misalignment (i.e. $\cos{\theta_i} < 0.0$) \citep{o3b_astro_dist,RouletGWTC2Pop,BuildBetterSpinModels,GWTC3MonashSpin,Callister_NoEvidence}. 
Another study recently showed that the spin tilt distribution inference is sensitive to modeling choices and may not peak at perfectly aligned spins, as is often assumed \citep{spinitasyoulike}. 
Other approaches to building improved models for BBH spins again, as with the mass models, focus on iterating through different simple parametric descriptions with 
increasing complexities \citep{BuildBetterSpinModels,Callister_NoEvidence,spinitasyoulike}. The spin distributions offer a nice use case for flexible data-driven models, as 
we present in this manuscript, both to help alleviate the current disputes in the literature and because there has been little work on flexible model development for the 
spin distribution of BBHs beyond assuming aligned spins thus far \citep{Tiwari_2021_a,Tiwari_2021_b}.

Polynomial splines have been applied to success across different areas of gravitational-wave astronomy. They have been used to model the gravitational-wave data noise spectrum, 
detector calibration uncertainties, coherent gravitational waveform deviations, and modulations to a power law mass distribution \citep{Littenberg_2015,Edwards_2018,B_Farr_etal_2014,Edelman_2021,Edelman_2022ApJ}
In this paper we highlight how the use of basis-splines can provide a powerful non-parametric modeling approach to the astrophysical distributions of compact 
binaries. We illustrate how one can efficiently model both the mass and spin distributions of merging compact binaries in GWTC-3 with basis splines to infer CBC population properties using 
hierarchical Bayesian inference. We discuss our results in the context of current literature on compact object populations and how this method complements the simpler lower 
dimensional parametric models in the short run, and will become necessary with future catalogs. This data-driven approach will provide checks of our understanding by finding more 
subtle features, that may not be expected, should they appear in with more observations. The rest of this manuscript is structured by first going over the background of basis splines in 
section \ref{sec:basis_splines}, followed by presenting the results of our extensive data-driven study of the mass and spin distributions of BBHs in GWTC-3 section \ref{sec:results}. We 
discuss our results and their astrophysical implications in section \ref{sec:astrodiscussion}, finishing with our conclusions in section \ref{sec:conclusion}.