### 2022-10-15 src/tex/macros.tex
add ppc plot and some tables to try and make results presentation cleaner and more complete

\makeatletter\newcommand\macros[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@out{\{"PLPeak": \{"beta": \{"median": 1.1, "error plus": 1.7, "error minus": 1.3, "5th percentile": {-}0.25, "95th percentile": 2.8, "10th percentile": 0.016, "90th percentile": 2.4\}, "local\_rate": \{"median": 17, "error plus": 9.6, "error minus": 6.7, "5th percentile": 11, "95th percentile": 28, "10th percentile": 12, "90th percentile": 25\}, "lamb": \{"median": 2.7, "error plus": 1.8, "error minus": 1.9, "5th percentile": 0.86, "95th percentile": 4.5, "10th percentile": 1.3, "90th percentile": 4.1\}\}, "Default": \{"a\_90percentile": \{"median": 0.53, "error plus": 0.094, "error minus": 0.073, "5th percentile": 0.46, "95th percentile": 0.63, "10th percentile": 0.47, "90th percentile": 0.6\}, "peak\_a": \{"median": 0.16, "error plus": 0.11, "error minus": 0.13, "5th percentile": 0.03, "95th percentile": 0.27, "10th percentile": 0.053, "90th percentile": 0.25\}, "lamb": \{"median": 2.7, "error plus": 1.8, "error minus": 1.9, "5th percentile": 0.86, "95th percentile": 4.5, "10th percentile": 1.3, "90th percentile": 4.1\}, "local\_rate": \{"median": 17, "error plus": 9.6, "error minus": 6.7, "5th percentile": 11, "95th percentile": 28, "10th percentile": 12, "90th percentile": 25\}, "peakCosTilt": \{"median": 1.0, "error plus": 0.0, "error minus": 0.0, "5th percentile": 1.0, "95th percentile": 1.0, "10th percentile": 1.0, "90th percentile": 1.0\}, "log10gammaFrac": \{"median": 0.19, "error plus": 0.36, "error minus": 0.17, "5th percentile": 0.022, "95th percentile": 0.59, "10th percentile": 0.039, "90th percentile": 0.5\}, "negFrac": \{"median": 0.44, "error plus": 0.064, "error minus": 0.12, "5th percentile": 0.32, "95th percentile": 0.49, "10th percentile": 0.35, "90th percentile": 0.49\}\}, "BSplineIndependentCompSpins": \{"lamb": \{"median": 2.3, "error plus": 2.9, "error minus": 3.0, "5th percentile": {-}0.63, "95th percentile": 5.2, "10th percentile": {-}0.024, "90th percentile": 4.7\}, "local\_rate": \{"median": 18, "error plus": 45, "error minus": 14, "5th percentile": 4.2, "95th percentile": 71, "10th percentile": 5.9, "90th percentile": 52\}, "a1\_90percentile": \{"median": 0.59, "error plus": 0.24, "error minus": 0.17, "5th percentile": 0.42, "95th percentile": 0.85, "10th percentile": 0.45, "90th percentile": 0.79\}, "a2\_90percentile": \{"median": 0.59, "error plus": 0.29, "error minus": 0.3, "5th percentile": 0.3, "95th percentile": 0.87, "10th percentile": 0.35, "90th percentile": 0.82\}, "peak\_a1": \{"median": 0.24, "error plus": 0.16, "error minus": 0.24, "5th percentile": 0.0, "95th percentile": 0.38, "10th percentile": 0.052, "90th percentile": 0.34\}, "peak\_a2": \{"median": 0.11, "error plus": 0.14, "error minus": 0.11, "5th percentile": 0.0, "95th percentile": 0.26, "10th percentile": 0.0, "90th percentile": 0.22\}, "peakCosTilt1": \{"median": 0.48, "error plus": 0.59, "error minus": 0.74, "5th percentile": {-}0.26, "95th percentile": 1.0, "10th percentile": {-}0.11, "90th percentile": 0.86\}, "peakCosTilt2": \{"median": 0.39, "error plus": 0.82, "error minus": 1.4, "5th percentile": {-}1.0, "95th percentile": 1.0, "10th percentile": {-}0.87, "90th percentile": 1.0\}, "log10gammaFrac1": \{"median": 0.47, "error plus": 0.84, "error minus": 0.84, "5th percentile": {-}0.37, "95th percentile": 1.3, "10th percentile": {-}0.19, "90th percentile": 1.1\}, "log10gammaFrac2": \{"median": 0.22, "error plus": 1.0, "error minus": 1.0, "5th percentile": {-}0.79, "95th percentile": 1.3, "10th percentile": {-}0.58, "90th percentile": 1.0\}, "negFrac1": \{"median": 0.26, "error plus": 0.29, "error minus": 0.17, "5th percentile": 0.09, "95th percentile": 0.58, "10th percentile": 0.11, "90th percentile": 0.51\}, "negFrac2": \{"median": 0.4, "error plus": 0.37, "error minus": 0.29, "5th percentile": 0.11, "95th percentile": 0.8, "10th percentile": 0.14, "90th percentile": 0.74\}\}, "BSplineIIDCompSpins": \{"a\_90percentile": \{"median": 0.51, "error plus": 0.23, "error minus": 0.14, "5th percentile": 0.36, "95th percentile": 0.76, "10th percentile": 0.38, "90th percentile": 0.7\}, "peak\_a": \{"median": 0.2, "error plus": 0.083, "error minus": 0.18, "5th percentile": 0.017, "95th percentile": 0.27, "10th percentile": 0.087, "90th percentile": 0.26\}, "lamb": \{"median": 2.4, "error plus": 2.8, "error minus": 3.1, "5th percentile": {-}0.65, "95th percentile": 5.3, "10th percentile": 0.15, "90th percentile": 4.7\}, "local\_rate": \{"median": 16, "error plus": 46, "error minus": 12, "5th percentile": 4.2, "95th percentile": 70, "10th percentile": 5.8, "90th percentile": 51\}, "peakCosTilt": \{"median": 0.44, "error plus": 0.51, "error minus": 0.44, "5th percentile": {-}0.0013, "95th percentile": 1.0, "10th percentile": 0.12, "90th percentile": 1.0\}, "log10gammaFrac": \{"median": 0.34, "error plus": 0.8, "error minus": 0.74, "5th percentile": {-}0.4, "95th percentile": 1.2, "10th percentile": {-}0.25, "90th percentile": 0.98\}, "negFrac": \{"median": 0.27, "error plus": 0.14, "error minus": 0.11, "5th percentile": 0.16, "95th percentile": 0.41, "10th percentile": 0.18, "90th percentile": 0.37\}\}, "ChiEffective": \{"default": \{"FracBelowNeg0p3": \{"median": 0.013, "error plus": 0.018, "error minus": 0.0091, "5th percentile": 0.0042, "95th percentile": 0.033, "10th percentile": 0.0053, "90th percentile": 0.027\}, "FracBelow0": \{"median": 0.43, "error plus": 0.072, "error minus": 0.13, "5th percentile": 0.3, "95th percentile": 0.49, "10th percentile": 0.34, "90th percentile": 0.48\}, "PeakChiEff": \{"median": 0.017, "error plus": 0.03, "error minus": 0.022, "5th percentile": {-}0.005, "95th percentile": 0.049, "10th percentile": {-}0.001, "90th percentile": 0.043\}, "frac\_dyn": \{"median": 0.86, "error plus": 0.14, "error minus": 0.26, "5th percentile": 0.6, "95th percentile": 0.98, "10th percentile": 0.68, "90th percentile": 0.97\}, "frac\_hm": \{"median": 0.083, "error plus": 0.11, "error minus": 0.057, "5th percentile": 0.026, "95th percentile": 0.21, "10th percentile": 0.033, "90th percentile": 0.17\}\}, "iid": \{"FracBelowNeg0p3": \{"median": 0.0073, "error plus": 0.012, "error minus": 0.0057, "5th percentile": 0.0016, "95th percentile": 0.021, "10th percentile": 0.0024, "90th percentile": 0.017\}, "FracBelow0": \{"median": 0.27, "error plus": 0.12, "error minus": 0.1, "5th percentile": 0.17, "95th percentile": 0.39, "10th percentile": 0.18, "90th percentile": 0.37\}, "PeakChiEff": \{"median": 0.053, "error plus": 0.034, "error minus": 0.04, "5th percentile": 0.013, "95th percentile": 0.085, "10th percentile": 0.023, "90th percentile": 0.075\}, "frac\_dyn": \{"median": 0.55, "error plus": 0.24, "error minus": 0.21, "5th percentile": 0.34, "95th percentile": 0.79, "10th percentile": 0.37, "90th percentile": 0.74\}, "frac\_hm": \{"median": 0.046, "error plus": 0.077, "error minus": 0.036, "5th percentile": 0.01, "95th percentile": 0.13, "10th percentile": 0.015, "90th percentile": 0.11\}\}, "ind": \{"FracBelowNeg0p3": \{"median": 0.011, "error plus": 0.015, "error minus": 0.0077, "5th percentile": 0.0029, "95th percentile": 0.028, "10th percentile": 0.004, "90th percentile": 0.023\}, "FracBelow0": \{"median": 0.31, "error plus": 0.1, "error minus": 0.11, "5th percentile": 0.2, "95th percentile": 0.41, "10th percentile": 0.22, "90th percentile": 0.39\}, "PeakChiEff": \{"median": 0.039, "error plus": 0.033, "error minus": 0.028, "5th percentile": 0.011, "95th percentile": 0.073, "10th percentile": 0.017, "90th percentile": 0.063\}, "frac\_dyn": \{"median": 0.61, "error plus": 0.2, "error minus": 0.22, "5th percentile": 0.39, "95th percentile": 0.81, "10th percentile": 0.44, "90th percentile": 0.77\}, "frac\_hm": \{"median": 0.066, "error plus": 0.096, "error minus": 0.048, "5th percentile": 0.018, "95th percentile": 0.18, "10th percentile": 0.025, "90th percentile": 0.14\}\}, "gaussian": \{"FracBelowNeg0p3": \{"median": 0.00024, "error plus": 0.0067, "error minus": 0.00024, "5th percentile": 1.1e{-}07, "95th percentile": 0.0083, "10th percentile": 9.2e{-}07, "90th percentile": 0.0039\}, "FracBelow0": \{"median": 0.28, "error plus": 0.14, "error minus": 0.13, "5th percentile": 0.15, "95th percentile": 0.42, "10th percentile": 0.18, "90th percentile": 0.39\}, "PeakChiEff": \{"median": 0.06, "error plus": 0.031, "error minus": 0.037, "5th percentile": 0.023, "95th percentile": 0.089, "10th percentile": 0.032, "90th percentile": 0.082\}, "frac\_dyn": \{"median": 0.55, "error plus": 0.29, "error minus": 0.26, "5th percentile": 0.3, "95th percentile": 0.85, "10th percentile": 0.35, "90th percentile": 0.78\}, "frac\_hm": \{"median": 0.0015, "error plus": 0.042, "error minus": 0.0015, "5th percentile": 6.9e{-}07, "95th percentile": 0.052, "10th percentile": 5.7e{-}06, "90th percentile": 0.024\}\}\}, "MassDistribution": \{"PLPeak": \{"m\_1percentile": \{"median": 6.2, "error plus": 0.64, "error minus": 1.1, "5th percentile": 5.1, "95th percentile": 6.8, "10th percentile": 5.4, "90th percentile": 6.7\}, "m\_75percentile": \{"median": 15, "error plus": 5.5, "error minus": 2.7, "5th percentile": 12, "95th percentile": 21, "10th percentile": 13, "90th percentile": 19\}, "m\_99percentile": \{"median": 45, "error plus": 9.9, "error minus": 6.2, "5th percentile": 39, "95th percentile": 55, "10th percentile": 40, "90th percentile": 52\}\}, "BSpline": \{"m\_1percentile": \{"median": 8.7, "error plus": 0.64, "error minus": 0.94, "5th percentile": 7.8, "95th percentile": 9.3, "10th percentile": 8.1, "90th percentile": 9.2\}, "m\_75percentile": \{"median": 12, "error plus": 13, "error minus": 1.1, "5th percentile": 10, "95th percentile": 27, "10th percentile": 11, "90th percentile": 21\}, "m\_99percentile": \{"median": 45, "error plus": 10, "error minus": 6.2, "5th percentile": 39, "95th percentile": 57, "10th percentile": 40, "90th percentile": 54\}\}, "PLSpline": \{"m\_1percentile": \{"median": 6.2, "error plus": 0.64, "error minus": 1.1, "5th percentile": 5.1, "95th percentile": 6.8, "10th percentile": 5.4, "90th percentile": 6.7\}, "m\_75percentile": \{"median": 11, "error plus": 4.6, "error minus": 0.59, "5th percentile": 11, "95th percentile": 17, "10th percentile": 11, "90th percentile": 15\}, "m\_99percentile": \{"median": 43, "error plus": 10, "error minus": 4.7, "5th percentile": 38, "95th percentile": 55, "10th percentile": 39, "90th percentile": 52\}\}\}\}}\else\ifnum\pdfstrcmp{#1}{PLPeak}=0\let\macros@out\macros@I\else\ifnum\pdfstrcmp{#1}{Default}=0\let\macros@out\macros@II\else\ifnum\pdfstrcmp{#1}{BSplineIndependentCompSpins}=0\let\macros@out\macros@III\else\ifnum\pdfstrcmp{#1}{BSplineIIDCompSpins}=0\let\macros@out\macros@IV\else\ifnum\pdfstrcmp{#1}{ChiEffective}=0\let\macros@out\macros@V\else\ifnum\pdfstrcmp{#1}{MassDistribution}=0\let\macros@out\macros@VI\else\def\macros@out{??}\fi\fi\fi\fi\fi\fi\fi\macros@out}\newcommand\macros@I[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@I@out{\{"beta": \{"median": 1.1, "error plus": 1.7, "error minus": 1.3, "5th percentile": {-}0.25, "95th percentile": 2.8, "10th percentile": 0.016, "90th percentile": 2.4\}, "local\_rate": \{"median": 17, "error plus": 9.6, "error minus": 6.7, "5th percentile": 11, "95th percentile": 28, "10th percentile": 12, "90th percentile": 25\}, "lamb": \{"median": 2.7, "error plus": 1.8, "error minus": 1.9, "5th percentile": 0.86, "95th percentile": 4.5, "10th percentile": 1.3, "90th percentile": 4.1\}\}}\else\ifnum\pdfstrcmp{#1}{beta}=0\let\macros@I@out\macros@VII\else\ifnum\pdfstrcmp{#1}{local_rate}=0\let\macros@I@out\macros@VIII\else\ifnum\pdfstrcmp{#1}{lamb}=0\let\macros@I@out\macros@IX\else\def\macros@I@out{??}\fi\fi\fi\fi\macros@I@out}\newcommand\macros@II[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@II@out{\{"a\_90percentile": \{"median": 0.53, "error plus": 0.094, "error minus": 0.073, "5th percentile": 0.46, "95th percentile": 0.63, "10th percentile": 0.47, "90th percentile": 0.6\}, "peak\_a": \{"median": 0.16, "error plus": 0.11, "error minus": 0.13, "5th percentile": 0.03, "95th percentile": 0.27, "10th percentile": 0.053, "90th percentile": 0.25\}, "lamb": \{"median": 2.7, "error plus": 1.8, "error minus": 1.9, "5th percentile": 0.86, "95th percentile": 4.5, "10th percentile": 1.3, "90th percentile": 4.1\}, "local\_rate": \{"median": 17, "error plus": 9.6, "error minus": 6.7, "5th percentile": 11, "95th percentile": 28, "10th percentile": 12, "90th percentile": 25\}, "peakCosTilt": \{"median": 1.0, "error plus": 0.0, "error minus": 0.0, "5th percentile": 1.0, "95th percentile": 1.0, "10th percentile": 1.0, "90th percentile": 1.0\}, "log10gammaFrac": \{"median": 0.19, "error plus": 0.36, "error minus": 0.17, "5th percentile": 0.022, "95th percentile": 0.59, "10th percentile": 0.039, "90th percentile": 0.5\}, "negFrac": \{"median": 0.44, "error plus": 0.064, "error minus": 0.12, "5th percentile": 0.32, "95th percentile": 0.49, "10th percentile": 0.35, "90th percentile": 0.49\}\}}\else\ifnum\pdfstrcmp{#1}{a_90percentile}=0\let\macros@II@out\macros@X\else\ifnum\pdfstrcmp{#1}{peak_a}=0\let\macros@II@out\macros@XI\else\ifnum\pdfstrcmp{#1}{lamb}=0\let\macros@II@out\macros@XII\else\ifnum\pdfstrcmp{#1}{local_rate}=0\let\macros@II@out\macros@XIII\else\ifnum\pdfstrcmp{#1}{peakCosTilt}=0\let\macros@II@out\macros@XIV\else\ifnum\pdfstrcmp{#1}{log10gammaFrac}=0\let\macros@II@out\macros@XV\else\ifnum\pdfstrcmp{#1}{negFrac}=0\let\macros@II@out\macros@XVI\else\def\macros@II@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@II@out}\newcommand\macros@III[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@III@out{\{"lamb": \{"median": 2.3, "error plus": 2.9, "error minus": 3.0, "5th percentile": {-}0.63, "95th percentile": 5.2, "10th percentile": {-}0.024, "90th percentile": 4.7\}, "local\_rate": \{"median": 18, "error plus": 45, "error minus": 14, "5th percentile": 4.2, "95th percentile": 71, "10th percentile": 5.9, "90th percentile": 52\}, "a1\_90percentile": \{"median": 0.59, "error plus": 0.24, "error minus": 0.17, "5th percentile": 0.42, "95th percentile": 0.85, "10th percentile": 0.45, "90th percentile": 0.79\}, "a2\_90percentile": \{"median": 0.59, "error plus": 0.29, "error minus": 0.3, "5th percentile": 0.3, "95th percentile": 0.87, "10th percentile": 0.35, "90th percentile": 0.82\}, "peak\_a1": \{"median": 0.24, "error plus": 0.16, "error minus": 0.24, "5th percentile": 0.0, "95th percentile": 0.38, "10th percentile": 0.052, "90th percentile": 0.34\}, "peak\_a2": \{"median": 0.11, "error plus": 0.14, "error minus": 0.11, "5th percentile": 0.0, "95th percentile": 0.26, "10th percentile": 0.0, "90th percentile": 0.22\}, "peakCosTilt1": \{"median": 0.48, "error plus": 0.59, "error minus": 0.74, "5th percentile": {-}0.26, "95th percentile": 1.0, "10th percentile": {-}0.11, "90th percentile": 0.86\}, "peakCosTilt2": \{"median": 0.39, "error plus": 0.82, "error minus": 1.4, "5th percentile": {-}1.0, "95th percentile": 1.0, "10th percentile": {-}0.87, "90th percentile": 1.0\}, "log10gammaFrac1": \{"median": 0.47, "error plus": 0.84, "error minus": 0.84, "5th percentile": {-}0.37, "95th percentile": 1.3, "10th percentile": {-}0.19, "90th percentile": 1.1\}, "log10gammaFrac2": \{"median": 0.22, "error plus": 1.0, "error minus": 1.0, "5th percentile": {-}0.79, "95th percentile": 1.3, "10th percentile": {-}0.58, "90th percentile": 1.0\}, "negFrac1": \{"median": 0.26, "error plus": 0.29, "error minus": 0.17, "5th percentile": 0.09, "95th percentile": 0.58, "10th percentile": 0.11, "90th percentile": 0.51\}, "negFrac2": \{"median": 0.4, "error plus": 0.37, "error minus": 0.29, "5th percentile": 0.11, "95th percentile": 0.8, "10th percentile": 0.14, "90th percentile": 0.74\}\}}\else\ifnum\pdfstrcmp{#1}{lamb}=0\let\macros@III@out\macros@XVII\else\ifnum\pdfstrcmp{#1}{local_rate}=0\let\macros@III@out\macros@XVIII\else\ifnum\pdfstrcmp{#1}{a1_90percentile}=0\let\macros@III@out\macros@XIX\else\ifnum\pdfstrcmp{#1}{a2_90percentile}=0\let\macros@III@out\macros@XX\else\ifnum\pdfstrcmp{#1}{peak_a1}=0\let\macros@III@out\macros@XXI\else\ifnum\pdfstrcmp{#1}{peak_a2}=0\let\macros@III@out\macros@XXII\else\ifnum\pdfstrcmp{#1}{peakCosTilt1}=0\let\macros@III@out\macros@XXIII\else\ifnum\pdfstrcmp{#1}{peakCosTilt2}=0\let\macros@III@out\macros@XXIV\else\ifnum\pdfstrcmp{#1}{log10gammaFrac1}=0\let\macros@III@out\macros@XXV\else\ifnum\pdfstrcmp{#1}{log10gammaFrac2}=0\let\macros@III@out\macros@XXVI\else\ifnum\pdfstrcmp{#1}{negFrac1}=0\let\macros@III@out\macros@XXVII\else\ifnum\pdfstrcmp{#1}{negFrac2}=0\let\macros@III@out\macros@XXVIII\else\def\macros@III@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\fi\fi\fi\fi\fi\macros@III@out}\newcommand\macros@IV[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@IV@out{\{"a\_90percentile": \{"median": 0.51, "error plus": 0.23, "error minus": 0.14, "5th percentile": 0.36, "95th percentile": 0.76, "10th percentile": 0.38, "90th percentile": 0.7\}, "peak\_a": \{"median": 0.2, "error plus": 0.083, "error minus": 0.18, "5th percentile": 0.017, "95th percentile": 0.27, "10th percentile": 0.087, "90th percentile": 0.26\}, "lamb": \{"median": 2.4, "error plus": 2.8, "error minus": 3.1, "5th percentile": {-}0.65, "95th percentile": 5.3, "10th percentile": 0.15, "90th percentile": 4.7\}, "local\_rate": \{"median": 16, "error plus": 46, "error minus": 12, "5th percentile": 4.2, "95th percentile": 70, "10th percentile": 5.8, "90th percentile": 51\}, "peakCosTilt": \{"median": 0.44, "error plus": 0.51, "error minus": 0.44, "5th percentile": {-}0.0013, "95th percentile": 1.0, "10th percentile": 0.12, "90th percentile": 1.0\}, "log10gammaFrac": \{"median": 0.34, "error plus": 0.8, "error minus": 0.74, "5th percentile": {-}0.4, "95th percentile": 1.2, "10th percentile": {-}0.25, "90th percentile": 0.98\}, "negFrac": \{"median": 0.27, "error plus": 0.14, "error minus": 0.11, "5th percentile": 0.16, "95th percentile": 0.41, "10th percentile": 0.18, "90th percentile": 0.37\}\}}\else\ifnum\pdfstrcmp{#1}{a_90percentile}=0\let\macros@IV@out\macros@XXIX\else\ifnum\pdfstrcmp{#1}{peak_a}=0\let\macros@IV@out\macros@XXX\else\ifnum\pdfstrcmp{#1}{lamb}=0\let\macros@IV@out\macros@XXXI\else\ifnum\pdfstrcmp{#1}{local_rate}=0\let\macros@IV@out\macros@XXXII\else\ifnum\pdfstrcmp{#1}{peakCosTilt}=0\let\macros@IV@out\macros@XXXIII\else\ifnum\pdfstrcmp{#1}{log10gammaFrac}=0\let\macros@IV@out\macros@XXXIV\else\ifnum\pdfstrcmp{#1}{negFrac}=0\let\macros@IV@out\macros@XXXV\else\def\macros@IV@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@IV@out}\newcommand\macros@V[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@V@out{\{"default": \{"FracBelowNeg0p3": \{"median": 0.013, "error plus": 0.018, "error minus": 0.0091, "5th percentile": 0.0042, "95th percentile": 0.033, "10th percentile": 0.0053, "90th percentile": 0.027\}, "FracBelow0": \{"median": 0.43, "error plus": 0.072, "error minus": 0.13, "5th percentile": 0.3, "95th percentile": 0.49, "10th percentile": 0.34, "90th percentile": 0.48\}, "PeakChiEff": \{"median": 0.017, "error plus": 0.03, "error minus": 0.022, "5th percentile": {-}0.005, "95th percentile": 0.049, "10th percentile": {-}0.001, "90th percentile": 0.043\}, "frac\_dyn": \{"median": 0.86, "error plus": 0.14, "error minus": 0.26, "5th percentile": 0.6, "95th percentile": 0.98, "10th percentile": 0.68, "90th percentile": 0.97\}, "frac\_hm": \{"median": 0.083, "error plus": 0.11, "error minus": 0.057, "5th percentile": 0.026, "95th percentile": 0.21, "10th percentile": 0.033, "90th percentile": 0.17\}\}, "iid": \{"FracBelowNeg0p3": \{"median": 0.0073, "error plus": 0.012, "error minus": 0.0057, "5th percentile": 0.0016, "95th percentile": 0.021, "10th percentile": 0.0024, "90th percentile": 0.017\}, "FracBelow0": \{"median": 0.27, "error plus": 0.12, "error minus": 0.1, "5th percentile": 0.17, "95th percentile": 0.39, "10th percentile": 0.18, "90th percentile": 0.37\}, "PeakChiEff": \{"median": 0.053, "error plus": 0.034, "error minus": 0.04, "5th percentile": 0.013, "95th percentile": 0.085, "10th percentile": 0.023, "90th percentile": 0.075\}, "frac\_dyn": \{"median": 0.55, "error plus": 0.24, "error minus": 0.21, "5th percentile": 0.34, "95th percentile": 0.79, "10th percentile": 0.37, "90th percentile": 0.74\}, "frac\_hm": \{"median": 0.046, "error plus": 0.077, "error minus": 0.036, "5th percentile": 0.01, "95th percentile": 0.13, "10th percentile": 0.015, "90th percentile": 0.11\}\}, "ind": \{"FracBelowNeg0p3": \{"median": 0.011, "error plus": 0.015, "error minus": 0.0077, "5th percentile": 0.0029, "95th percentile": 0.028, "10th percentile": 0.004, "90th percentile": 0.023\}, "FracBelow0": \{"median": 0.31, "error plus": 0.1, "error minus": 0.11, "5th percentile": 0.2, "95th percentile": 0.41, "10th percentile": 0.22, "90th percentile": 0.39\}, "PeakChiEff": \{"median": 0.039, "error plus": 0.033, "error minus": 0.028, "5th percentile": 0.011, "95th percentile": 0.073, "10th percentile": 0.017, "90th percentile": 0.063\}, "frac\_dyn": \{"median": 0.61, "error plus": 0.2, "error minus": 0.22, "5th percentile": 0.39, "95th percentile": 0.81, "10th percentile": 0.44, "90th percentile": 0.77\}, "frac\_hm": \{"median": 0.066, "error plus": 0.096, "error minus": 0.048, "5th percentile": 0.018, "95th percentile": 0.18, "10th percentile": 0.025, "90th percentile": 0.14\}\}, "gaussian": \{"FracBelowNeg0p3": \{"median": 0.00024, "error plus": 0.0067, "error minus": 0.00024, "5th percentile": 1.1e{-}07, "95th percentile": 0.0083, "10th percentile": 9.2e{-}07, "90th percentile": 0.0039\}, "FracBelow0": \{"median": 0.28, "error plus": 0.14, "error minus": 0.13, "5th percentile": 0.15, "95th percentile": 0.42, "10th percentile": 0.18, "90th percentile": 0.39\}, "PeakChiEff": \{"median": 0.06, "error plus": 0.031, "error minus": 0.037, "5th percentile": 0.023, "95th percentile": 0.089, "10th percentile": 0.032, "90th percentile": 0.082\}, "frac\_dyn": \{"median": 0.55, "error plus": 0.29, "error minus": 0.26, "5th percentile": 0.3, "95th percentile": 0.85, "10th percentile": 0.35, "90th percentile": 0.78\}, "frac\_hm": \{"median": 0.0015, "error plus": 0.042, "error minus": 0.0015, "5th percentile": 6.9e{-}07, "95th percentile": 0.052, "10th percentile": 5.7e{-}06, "90th percentile": 0.024\}\}\}}\else\ifnum\pdfstrcmp{#1}{default}=0\let\macros@V@out\macros@XXXVI\else\ifnum\pdfstrcmp{#1}{iid}=0\let\macros@V@out\macros@XXXVII\else\ifnum\pdfstrcmp{#1}{ind}=0\let\macros@V@out\macros@XXXVIII\else\ifnum\pdfstrcmp{#1}{gaussian}=0\let\macros@V@out\macros@XXXIX\else\def\macros@V@out{??}\fi\fi\fi\fi\fi\macros@V@out}\newcommand\macros@VI[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@VI@out{\{"PLPeak": \{"m\_1percentile": \{"median": 6.2, "error plus": 0.64, "error minus": 1.1, "5th percentile": 5.1, "95th percentile": 6.8, "10th percentile": 5.4, "90th percentile": 6.7\}, "m\_75percentile": \{"median": 15, "error plus": 5.5, "error minus": 2.7, "5th percentile": 12, "95th percentile": 21, "10th percentile": 13, "90th percentile": 19\}, "m\_99percentile": \{"median": 45, "error plus": 9.9, "error minus": 6.2, "5th percentile": 39, "95th percentile": 55, "10th percentile": 40, "90th percentile": 52\}\}, "BSpline": \{"m\_1percentile": \{"median": 8.7, "error plus": 0.64, "error minus": 0.94, "5th percentile": 7.8, "95th percentile": 9.3, "10th percentile": 8.1, "90th percentile": 9.2\}, "m\_75percentile": \{"median": 12, "error plus": 13, "error minus": 1.1, "5th percentile": 10, "95th percentile": 27, "10th percentile": 11, "90th percentile": 21\}, "m\_99percentile": \{"median": 45, "error plus": 10, "error minus": 6.2, "5th percentile": 39, "95th percentile": 57, "10th percentile": 40, "90th percentile": 54\}\}, "PLSpline": \{"m\_1percentile": \{"median": 6.2, "error plus": 0.64, "error minus": 1.1, "5th percentile": 5.1, "95th percentile": 6.8, "10th percentile": 5.4, "90th percentile": 6.7\}, "m\_75percentile": \{"median": 11, "error plus": 4.6, "error minus": 0.59, "5th percentile": 11, "95th percentile": 17, "10th percentile": 11, "90th percentile": 15\}, "m\_99percentile": \{"median": 43, "error plus": 10, "error minus": 4.7, "5th percentile": 38, "95th percentile": 55, "10th percentile": 39, "90th percentile": 52\}\}\}}\else\ifnum\pdfstrcmp{#1}{PLPeak}=0\let\macros@VI@out\macros@XL\else\ifnum\pdfstrcmp{#1}{BSpline}=0\let\macros@VI@out\macros@XLI\else\ifnum\pdfstrcmp{#1}{PLSpline}=0\let\macros@VI@out\macros@XLII\else\def\macros@VI@out{??}\fi\fi\fi\fi\macros@VI@out}\newcommand\macros@VII[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@VII@out{\{"median": 1.1, "error plus": 1.7, "error minus": 1.3, "5th percentile": {-}0.25, "95th percentile": 2.8, "10th percentile": 0.016, "90th percentile": 2.4\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@VII@out{1.1}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@VII@out{1.7}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@VII@out{1.3}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@VII@out{{-}0.25}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@VII@out{2.8}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@VII@out{0.016}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@VII@out{2.4}\else\def\macros@VII@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@VII@out}\newcommand\macros@VIII[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@VIII@out{\{"median": 17, "error plus": 9.6, "error minus": 6.7, "5th percentile": 11, "95th percentile": 28, "10th percentile": 12, "90th percentile": 25\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@VIII@out{17}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@VIII@out{9.6}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@VIII@out{6.7}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@VIII@out{11}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@VIII@out{28}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@VIII@out{12}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@VIII@out{25}\else\def\macros@VIII@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@VIII@out}\newcommand\macros@IX[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@IX@out{\{"median": 2.7, "error plus": 1.8, "error minus": 1.9, "5th percentile": 0.86, "95th percentile": 4.5, "10th percentile": 1.3, "90th percentile": 4.1\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@IX@out{2.7}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@IX@out{1.8}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@IX@out{1.9}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@IX@out{0.86}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@IX@out{4.5}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@IX@out{1.3}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@IX@out{4.1}\else\def\macros@IX@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@IX@out}\newcommand\macros@X[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@X@out{\{"median": 0.53, "error plus": 0.094, "error minus": 0.073, "5th percentile": 0.46, "95th percentile": 0.63, "10th percentile": 0.47, "90th percentile": 0.6\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@X@out{0.53}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@X@out{0.094}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@X@out{0.073}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@X@out{0.46}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@X@out{0.63}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@X@out{0.47}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@X@out{0.6}\else\def\macros@X@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@X@out}\newcommand\macros@XI[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XI@out{\{"median": 0.16, "error plus": 0.11, "error minus": 0.13, "5th percentile": 0.03, "95th percentile": 0.27, "10th percentile": 0.053, "90th percentile": 0.25\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XI@out{0.16}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XI@out{0.11}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XI@out{0.13}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XI@out{0.03}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XI@out{0.27}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XI@out{0.053}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XI@out{0.25}\else\def\macros@XI@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XI@out}\newcommand\macros@XII[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XII@out{\{"median": 2.7, "error plus": 1.8, "error minus": 1.9, "5th percentile": 0.86, "95th percentile": 4.5, "10th percentile": 1.3, "90th percentile": 4.1\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XII@out{2.7}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XII@out{1.8}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XII@out{1.9}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XII@out{0.86}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XII@out{4.5}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XII@out{1.3}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XII@out{4.1}\else\def\macros@XII@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XII@out}\newcommand\macros@XIII[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XIII@out{\{"median": 17, "error plus": 9.6, "error minus": 6.7, "5th percentile": 11, "95th percentile": 28, "10th percentile": 12, "90th percentile": 25\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XIII@out{17}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XIII@out{9.6}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XIII@out{6.7}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XIII@out{11}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XIII@out{28}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XIII@out{12}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XIII@out{25}\else\def\macros@XIII@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XIII@out}\newcommand\macros@XIV[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XIV@out{\{"median": 1.0, "error plus": 0.0, "error minus": 0.0, "5th percentile": 1.0, "95th percentile": 1.0, "10th percentile": 1.0, "90th percentile": 1.0\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XIV@out{1.0}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XIV@out{0.0}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XIV@out{0.0}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XIV@out{1.0}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XIV@out{1.0}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XIV@out{1.0}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XIV@out{1.0}\else\def\macros@XIV@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XIV@out}\newcommand\macros@XV[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XV@out{\{"median": 0.19, "error plus": 0.36, "error minus": 0.17, "5th percentile": 0.022, "95th percentile": 0.59, "10th percentile": 0.039, "90th percentile": 0.5\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XV@out{0.19}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XV@out{0.36}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XV@out{0.17}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XV@out{0.022}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XV@out{0.59}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XV@out{0.039}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XV@out{0.5}\else\def\macros@XV@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XV@out}\newcommand\macros@XVI[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XVI@out{\{"median": 0.44, "error plus": 0.064, "error minus": 0.12, "5th percentile": 0.32, "95th percentile": 0.49, "10th percentile": 0.35, "90th percentile": 0.49\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XVI@out{0.44}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XVI@out{0.064}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XVI@out{0.12}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XVI@out{0.32}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XVI@out{0.49}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XVI@out{0.35}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XVI@out{0.49}\else\def\macros@XVI@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XVI@out}\newcommand\macros@XVII[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XVII@out{\{"median": 2.3, "error plus": 2.9, "error minus": 3.0, "5th percentile": {-}0.63, "95th percentile": 5.2, "10th percentile": {-}0.024, "90th percentile": 4.7\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XVII@out{2.3}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XVII@out{2.9}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XVII@out{3.0}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XVII@out{{-}0.63}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XVII@out{5.2}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XVII@out{{-}0.024}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XVII@out{4.7}\else\def\macros@XVII@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XVII@out}\newcommand\macros@XVIII[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XVIII@out{\{"median": 18, "error plus": 45, "error minus": 14, "5th percentile": 4.2, "95th percentile": 71, "10th percentile": 5.9, "90th percentile": 52\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XVIII@out{18}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XVIII@out{45}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XVIII@out{14}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XVIII@out{4.2}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XVIII@out{71}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XVIII@out{5.9}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XVIII@out{52}\else\def\macros@XVIII@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XVIII@out}\newcommand\macros@XIX[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XIX@out{\{"median": 0.59, "error plus": 0.24, "error minus": 0.17, "5th percentile": 0.42, "95th percentile": 0.85, "10th percentile": 0.45, "90th percentile": 0.79\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XIX@out{0.59}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XIX@out{0.24}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XIX@out{0.17}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XIX@out{0.42}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XIX@out{0.85}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XIX@out{0.45}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XIX@out{0.79}\else\def\macros@XIX@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XIX@out}\newcommand\macros@XX[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XX@out{\{"median": 0.59, "error plus": 0.29, "error minus": 0.3, "5th percentile": 0.3, "95th percentile": 0.87, "10th percentile": 0.35, "90th percentile": 0.82\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XX@out{0.59}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XX@out{0.29}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XX@out{0.3}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XX@out{0.3}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XX@out{0.87}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XX@out{0.35}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XX@out{0.82}\else\def\macros@XX@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XX@out}\newcommand\macros@XXI[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XXI@out{\{"median": 0.24, "error plus": 0.16, "error minus": 0.24, "5th percentile": 0.0, "95th percentile": 0.38, "10th percentile": 0.052, "90th percentile": 0.34\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XXI@out{0.24}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XXI@out{0.16}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XXI@out{0.24}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XXI@out{0.0}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XXI@out{0.38}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XXI@out{0.052}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XXI@out{0.34}\else\def\macros@XXI@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XXI@out}\newcommand\macros@XXII[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XXII@out{\{"median": 0.11, "error plus": 0.14, "error minus": 0.11, "5th percentile": 0.0, "95th percentile": 0.26, "10th percentile": 0.0, "90th percentile": 0.22\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XXII@out{0.11}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XXII@out{0.14}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XXII@out{0.11}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XXII@out{0.0}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XXII@out{0.26}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XXII@out{0.0}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XXII@out{0.22}\else\def\macros@XXII@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XXII@out}\newcommand\macros@XXIII[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XXIII@out{\{"median": 0.48, "error plus": 0.59, "error minus": 0.74, "5th percentile": {-}0.26, "95th percentile": 1.0, "10th percentile": {-}0.11, "90th percentile": 0.86\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XXIII@out{0.48}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XXIII@out{0.59}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XXIII@out{0.74}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XXIII@out{{-}0.26}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XXIII@out{1.0}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XXIII@out{{-}0.11}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XXIII@out{0.86}\else\def\macros@XXIII@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XXIII@out}\newcommand\macros@XXIV[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XXIV@out{\{"median": 0.39, "error plus": 0.82, "error minus": 1.4, "5th percentile": {-}1.0, "95th percentile": 1.0, "10th percentile": {-}0.87, "90th percentile": 1.0\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XXIV@out{0.39}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XXIV@out{0.82}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XXIV@out{1.4}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XXIV@out{{-}1.0}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XXIV@out{1.0}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XXIV@out{{-}0.87}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XXIV@out{1.0}\else\def\macros@XXIV@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XXIV@out}\newcommand\macros@XXV[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XXV@out{\{"median": 0.47, "error plus": 0.84, "error minus": 0.84, "5th percentile": {-}0.37, "95th percentile": 1.3, "10th percentile": {-}0.19, "90th percentile": 1.1\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XXV@out{0.47}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XXV@out{0.84}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XXV@out{0.84}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XXV@out{{-}0.37}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XXV@out{1.3}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XXV@out{{-}0.19}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XXV@out{1.1}\else\def\macros@XXV@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XXV@out}\newcommand\macros@XXVI[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XXVI@out{\{"median": 0.22, "error plus": 1.0, "error minus": 1.0, "5th percentile": {-}0.79, "95th percentile": 1.3, "10th percentile": {-}0.58, "90th percentile": 1.0\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XXVI@out{0.22}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XXVI@out{1.0}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XXVI@out{1.0}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XXVI@out{{-}0.79}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XXVI@out{1.3}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XXVI@out{{-}0.58}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XXVI@out{1.0}\else\def\macros@XXVI@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XXVI@out}\newcommand\macros@XXVII[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XXVII@out{\{"median": 0.26, "error plus": 0.29, "error minus": 0.17, "5th percentile": 0.09, "95th percentile": 0.58, "10th percentile": 0.11, "90th percentile": 0.51\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XXVII@out{0.26}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XXVII@out{0.29}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XXVII@out{0.17}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XXVII@out{0.09}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XXVII@out{0.58}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XXVII@out{0.11}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XXVII@out{0.51}\else\def\macros@XXVII@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XXVII@out}\newcommand\macros@XXVIII[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XXVIII@out{\{"median": 0.4, "error plus": 0.37, "error minus": 0.29, "5th percentile": 0.11, "95th percentile": 0.8, "10th percentile": 0.14, "90th percentile": 0.74\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XXVIII@out{0.4}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XXVIII@out{0.37}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XXVIII@out{0.29}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XXVIII@out{0.11}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XXVIII@out{0.8}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XXVIII@out{0.14}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XXVIII@out{0.74}\else\def\macros@XXVIII@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XXVIII@out}\newcommand\macros@XXIX[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XXIX@out{\{"median": 0.51, "error plus": 0.23, "error minus": 0.14, "5th percentile": 0.36, "95th percentile": 0.76, "10th percentile": 0.38, "90th percentile": 0.7\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XXIX@out{0.51}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XXIX@out{0.23}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XXIX@out{0.14}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XXIX@out{0.36}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XXIX@out{0.76}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XXIX@out{0.38}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XXIX@out{0.7}\else\def\macros@XXIX@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XXIX@out}\newcommand\macros@XXX[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XXX@out{\{"median": 0.2, "error plus": 0.083, "error minus": 0.18, "5th percentile": 0.017, "95th percentile": 0.27, "10th percentile": 0.087, "90th percentile": 0.26\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XXX@out{0.2}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XXX@out{0.083}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XXX@out{0.18}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XXX@out{0.017}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XXX@out{0.27}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XXX@out{0.087}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XXX@out{0.26}\else\def\macros@XXX@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XXX@out}\newcommand\macros@XXXI[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XXXI@out{\{"median": 2.4, "error plus": 2.8, "error minus": 3.1, "5th percentile": {-}0.65, "95th percentile": 5.3, "10th percentile": 0.15, "90th percentile": 4.7\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XXXI@out{2.4}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XXXI@out{2.8}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XXXI@out{3.1}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XXXI@out{{-}0.65}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XXXI@out{5.3}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XXXI@out{0.15}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XXXI@out{4.7}\else\def\macros@XXXI@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XXXI@out}\newcommand\macros@XXXII[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XXXII@out{\{"median": 16, "error plus": 46, "error minus": 12, "5th percentile": 4.2, "95th percentile": 70, "10th percentile": 5.8, "90th percentile": 51\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XXXII@out{16}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XXXII@out{46}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XXXII@out{12}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XXXII@out{4.2}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XXXII@out{70}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XXXII@out{5.8}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XXXII@out{51}\else\def\macros@XXXII@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XXXII@out}\newcommand\macros@XXXIII[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XXXIII@out{\{"median": 0.44, "error plus": 0.51, "error minus": 0.44, "5th percentile": {-}0.0013, "95th percentile": 1.0, "10th percentile": 0.12, "90th percentile": 1.0\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XXXIII@out{0.44}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XXXIII@out{0.51}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XXXIII@out{0.44}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XXXIII@out{{-}0.0013}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XXXIII@out{1.0}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XXXIII@out{0.12}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XXXIII@out{1.0}\else\def\macros@XXXIII@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XXXIII@out}\newcommand\macros@XXXIV[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XXXIV@out{\{"median": 0.34, "error plus": 0.8, "error minus": 0.74, "5th percentile": {-}0.4, "95th percentile": 1.2, "10th percentile": {-}0.25, "90th percentile": 0.98\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XXXIV@out{0.34}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XXXIV@out{0.8}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XXXIV@out{0.74}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XXXIV@out{{-}0.4}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XXXIV@out{1.2}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XXXIV@out{{-}0.25}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XXXIV@out{0.98}\else\def\macros@XXXIV@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XXXIV@out}\newcommand\macros@XXXV[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XXXV@out{\{"median": 0.27, "error plus": 0.14, "error minus": 0.11, "5th percentile": 0.16, "95th percentile": 0.41, "10th percentile": 0.18, "90th percentile": 0.37\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XXXV@out{0.27}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XXXV@out{0.14}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XXXV@out{0.11}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XXXV@out{0.16}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XXXV@out{0.41}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XXXV@out{0.18}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XXXV@out{0.37}\else\def\macros@XXXV@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XXXV@out}\newcommand\macros@XXXVI[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XXXVI@out{\{"FracBelowNeg0p3": \{"median": 0.013, "error plus": 0.018, "error minus": 0.0091, "5th percentile": 0.0042, "95th percentile": 0.033, "10th percentile": 0.0053, "90th percentile": 0.027\}, "FracBelow0": \{"median": 0.43, "error plus": 0.072, "error minus": 0.13, "5th percentile": 0.3, "95th percentile": 0.49, "10th percentile": 0.34, "90th percentile": 0.48\}, "PeakChiEff": \{"median": 0.017, "error plus": 0.03, "error minus": 0.022, "5th percentile": {-}0.005, "95th percentile": 0.049, "10th percentile": {-}0.001, "90th percentile": 0.043\}, "frac\_dyn": \{"median": 0.86, "error plus": 0.14, "error minus": 0.26, "5th percentile": 0.6, "95th percentile": 0.98, "10th percentile": 0.68, "90th percentile": 0.97\}, "frac\_hm": \{"median": 0.083, "error plus": 0.11, "error minus": 0.057, "5th percentile": 0.026, "95th percentile": 0.21, "10th percentile": 0.033, "90th percentile": 0.17\}\}}\else\ifnum\pdfstrcmp{#1}{FracBelowNeg0p3}=0\let\macros@XXXVI@out\macros@XLIII\else\ifnum\pdfstrcmp{#1}{FracBelow0}=0\let\macros@XXXVI@out\macros@XLIV\else\ifnum\pdfstrcmp{#1}{PeakChiEff}=0\let\macros@XXXVI@out\macros@XLV\else\ifnum\pdfstrcmp{#1}{frac_dyn}=0\let\macros@XXXVI@out\macros@XLVI\else\ifnum\pdfstrcmp{#1}{frac_hm}=0\let\macros@XXXVI@out\macros@XLVII\else\def\macros@XXXVI@out{??}\fi\fi\fi\fi\fi\fi\macros@XXXVI@out}\newcommand\macros@XXXVII[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XXXVII@out{\{"FracBelowNeg0p3": \{"median": 0.0073, "error plus": 0.012, "error minus": 0.0057, "5th percentile": 0.0016, "95th percentile": 0.021, "10th percentile": 0.0024, "90th percentile": 0.017\}, "FracBelow0": \{"median": 0.27, "error plus": 0.12, "error minus": 0.1, "5th percentile": 0.17, "95th percentile": 0.39, "10th percentile": 0.18, "90th percentile": 0.37\}, "PeakChiEff": \{"median": 0.053, "error plus": 0.034, "error minus": 0.04, "5th percentile": 0.013, "95th percentile": 0.085, "10th percentile": 0.023, "90th percentile": 0.075\}, "frac\_dyn": \{"median": 0.55, "error plus": 0.24, "error minus": 0.21, "5th percentile": 0.34, "95th percentile": 0.79, "10th percentile": 0.37, "90th percentile": 0.74\}, "frac\_hm": \{"median": 0.046, "error plus": 0.077, "error minus": 0.036, "5th percentile": 0.01, "95th percentile": 0.13, "10th percentile": 0.015, "90th percentile": 0.11\}\}}\else\ifnum\pdfstrcmp{#1}{FracBelowNeg0p3}=0\let\macros@XXXVII@out\macros@XLVIII\else\ifnum\pdfstrcmp{#1}{FracBelow0}=0\let\macros@XXXVII@out\macros@XLIX\else\ifnum\pdfstrcmp{#1}{PeakChiEff}=0\let\macros@XXXVII@out\macros@L\else\ifnum\pdfstrcmp{#1}{frac_dyn}=0\let\macros@XXXVII@out\macros@LI\else\ifnum\pdfstrcmp{#1}{frac_hm}=0\let\macros@XXXVII@out\macros@LII\else\def\macros@XXXVII@out{??}\fi\fi\fi\fi\fi\fi\macros@XXXVII@out}\newcommand\macros@XXXVIII[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XXXVIII@out{\{"FracBelowNeg0p3": \{"median": 0.011, "error plus": 0.015, "error minus": 0.0077, "5th percentile": 0.0029, "95th percentile": 0.028, "10th percentile": 0.004, "90th percentile": 0.023\}, "FracBelow0": \{"median": 0.31, "error plus": 0.1, "error minus": 0.11, "5th percentile": 0.2, "95th percentile": 0.41, "10th percentile": 0.22, "90th percentile": 0.39\}, "PeakChiEff": \{"median": 0.039, "error plus": 0.033, "error minus": 0.028, "5th percentile": 0.011, "95th percentile": 0.073, "10th percentile": 0.017, "90th percentile": 0.063\}, "frac\_dyn": \{"median": 0.61, "error plus": 0.2, "error minus": 0.22, "5th percentile": 0.39, "95th percentile": 0.81, "10th percentile": 0.44, "90th percentile": 0.77\}, "frac\_hm": \{"median": 0.066, "error plus": 0.096, "error minus": 0.048, "5th percentile": 0.018, "95th percentile": 0.18, "10th percentile": 0.025, "90th percentile": 0.14\}\}}\else\ifnum\pdfstrcmp{#1}{FracBelowNeg0p3}=0\let\macros@XXXVIII@out\macros@LIII\else\ifnum\pdfstrcmp{#1}{FracBelow0}=0\let\macros@XXXVIII@out\macros@LIV\else\ifnum\pdfstrcmp{#1}{PeakChiEff}=0\let\macros@XXXVIII@out\macros@LV\else\ifnum\pdfstrcmp{#1}{frac_dyn}=0\let\macros@XXXVIII@out\macros@LVI\else\ifnum\pdfstrcmp{#1}{frac_hm}=0\let\macros@XXXVIII@out\macros@LVII\else\def\macros@XXXVIII@out{??}\fi\fi\fi\fi\fi\fi\macros@XXXVIII@out}\newcommand\macros@XXXIX[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XXXIX@out{\{"FracBelowNeg0p3": \{"median": 0.00024, "error plus": 0.0067, "error minus": 0.00024, "5th percentile": 1.1e{-}07, "95th percentile": 0.0083, "10th percentile": 9.2e{-}07, "90th percentile": 0.0039\}, "FracBelow0": \{"median": 0.28, "error plus": 0.14, "error minus": 0.13, "5th percentile": 0.15, "95th percentile": 0.42, "10th percentile": 0.18, "90th percentile": 0.39\}, "PeakChiEff": \{"median": 0.06, "error plus": 0.031, "error minus": 0.037, "5th percentile": 0.023, "95th percentile": 0.089, "10th percentile": 0.032, "90th percentile": 0.082\}, "frac\_dyn": \{"median": 0.55, "error plus": 0.29, "error minus": 0.26, "5th percentile": 0.3, "95th percentile": 0.85, "10th percentile": 0.35, "90th percentile": 0.78\}, "frac\_hm": \{"median": 0.0015, "error plus": 0.042, "error minus": 0.0015, "5th percentile": 6.9e{-}07, "95th percentile": 0.052, "10th percentile": 5.7e{-}06, "90th percentile": 0.024\}\}}\else\ifnum\pdfstrcmp{#1}{FracBelowNeg0p3}=0\let\macros@XXXIX@out\macros@LVIII\else\ifnum\pdfstrcmp{#1}{FracBelow0}=0\let\macros@XXXIX@out\macros@LIX\else\ifnum\pdfstrcmp{#1}{PeakChiEff}=0\let\macros@XXXIX@out\macros@LX\else\ifnum\pdfstrcmp{#1}{frac_dyn}=0\let\macros@XXXIX@out\macros@LXI\else\ifnum\pdfstrcmp{#1}{frac_hm}=0\let\macros@XXXIX@out\macros@LXII\else\def\macros@XXXIX@out{??}\fi\fi\fi\fi\fi\fi\macros@XXXIX@out}\newcommand\macros@XL[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XL@out{\{"m\_1percentile": \{"median": 6.2, "error plus": 0.64, "error minus": 1.1, "5th percentile": 5.1, "95th percentile": 6.8, "10th percentile": 5.4, "90th percentile": 6.7\}, "m\_75percentile": \{"median": 15, "error plus": 5.5, "error minus": 2.7, "5th percentile": 12, "95th percentile": 21, "10th percentile": 13, "90th percentile": 19\}, "m\_99percentile": \{"median": 45, "error plus": 9.9, "error minus": 6.2, "5th percentile": 39, "95th percentile": 55, "10th percentile": 40, "90th percentile": 52\}\}}\else\ifnum\pdfstrcmp{#1}{m_1percentile}=0\let\macros@XL@out\macros@LXIII\else\ifnum\pdfstrcmp{#1}{m_75percentile}=0\let\macros@XL@out\macros@LXIV\else\ifnum\pdfstrcmp{#1}{m_99percentile}=0\let\macros@XL@out\macros@LXV\else\def\macros@XL@out{??}\fi\fi\fi\fi\macros@XL@out}\newcommand\macros@XLI[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XLI@out{\{"m\_1percentile": \{"median": 8.7, "error plus": 0.64, "error minus": 0.94, "5th percentile": 7.8, "95th percentile": 9.3, "10th percentile": 8.1, "90th percentile": 9.2\}, "m\_75percentile": \{"median": 12, "error plus": 13, "error minus": 1.1, "5th percentile": 10, "95th percentile": 27, "10th percentile": 11, "90th percentile": 21\}, "m\_99percentile": \{"median": 45, "error plus": 10, "error minus": 6.2, "5th percentile": 39, "95th percentile": 57, "10th percentile": 40, "90th percentile": 54\}\}}\else\ifnum\pdfstrcmp{#1}{m_1percentile}=0\let\macros@XLI@out\macros@LXVI\else\ifnum\pdfstrcmp{#1}{m_75percentile}=0\let\macros@XLI@out\macros@LXVII\else\ifnum\pdfstrcmp{#1}{m_99percentile}=0\let\macros@XLI@out\macros@LXVIII\else\def\macros@XLI@out{??}\fi\fi\fi\fi\macros@XLI@out}\newcommand\macros@XLII[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XLII@out{\{"m\_1percentile": \{"median": 6.2, "error plus": 0.64, "error minus": 1.1, "5th percentile": 5.1, "95th percentile": 6.8, "10th percentile": 5.4, "90th percentile": 6.7\}, "m\_75percentile": \{"median": 11, "error plus": 4.6, "error minus": 0.59, "5th percentile": 11, "95th percentile": 17, "10th percentile": 11, "90th percentile": 15\}, "m\_99percentile": \{"median": 43, "error plus": 10, "error minus": 4.7, "5th percentile": 38, "95th percentile": 55, "10th percentile": 39, "90th percentile": 52\}\}}\else\ifnum\pdfstrcmp{#1}{m_1percentile}=0\let\macros@XLII@out\macros@LXIX\else\ifnum\pdfstrcmp{#1}{m_75percentile}=0\let\macros@XLII@out\macros@LXX\else\ifnum\pdfstrcmp{#1}{m_99percentile}=0\let\macros@XLII@out\macros@LXXI\else\def\macros@XLII@out{??}\fi\fi\fi\fi\macros@XLII@out}\newcommand\macros@XLIII[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XLIII@out{\{"median": 0.013, "error plus": 0.018, "error minus": 0.0091, "5th percentile": 0.0042, "95th percentile": 0.033, "10th percentile": 0.0053, "90th percentile": 0.027\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XLIII@out{0.013}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XLIII@out{0.018}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XLIII@out{0.0091}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XLIII@out{0.0042}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XLIII@out{0.033}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XLIII@out{0.0053}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XLIII@out{0.027}\else\def\macros@XLIII@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XLIII@out}\newcommand\macros@XLIV[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XLIV@out{\{"median": 0.43, "error plus": 0.072, "error minus": 0.13, "5th percentile": 0.3, "95th percentile": 0.49, "10th percentile": 0.34, "90th percentile": 0.48\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XLIV@out{0.43}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XLIV@out{0.072}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XLIV@out{0.13}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XLIV@out{0.3}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XLIV@out{0.49}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XLIV@out{0.34}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XLIV@out{0.48}\else\def\macros@XLIV@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XLIV@out}\newcommand\macros@XLV[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XLV@out{\{"median": 0.017, "error plus": 0.03, "error minus": 0.022, "5th percentile": {-}0.005, "95th percentile": 0.049, "10th percentile": {-}0.001, "90th percentile": 0.043\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XLV@out{0.017}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XLV@out{0.03}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XLV@out{0.022}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XLV@out{{-}0.005}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XLV@out{0.049}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XLV@out{{-}0.001}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XLV@out{0.043}\else\def\macros@XLV@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XLV@out}\newcommand\macros@XLVI[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XLVI@out{\{"median": 0.86, "error plus": 0.14, "error minus": 0.26, "5th percentile": 0.6, "95th percentile": 0.98, "10th percentile": 0.68, "90th percentile": 0.97\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XLVI@out{0.86}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XLVI@out{0.14}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XLVI@out{0.26}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XLVI@out{0.6}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XLVI@out{0.98}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XLVI@out{0.68}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XLVI@out{0.97}\else\def\macros@XLVI@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XLVI@out}\newcommand\macros@XLVII[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XLVII@out{\{"median": 0.083, "error plus": 0.11, "error minus": 0.057, "5th percentile": 0.026, "95th percentile": 0.21, "10th percentile": 0.033, "90th percentile": 0.17\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XLVII@out{0.083}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XLVII@out{0.11}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XLVII@out{0.057}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XLVII@out{0.026}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XLVII@out{0.21}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XLVII@out{0.033}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XLVII@out{0.17}\else\def\macros@XLVII@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XLVII@out}\newcommand\macros@XLVIII[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XLVIII@out{\{"median": 0.0073, "error plus": 0.012, "error minus": 0.0057, "5th percentile": 0.0016, "95th percentile": 0.021, "10th percentile": 0.0024, "90th percentile": 0.017\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XLVIII@out{0.0073}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XLVIII@out{0.012}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XLVIII@out{0.0057}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XLVIII@out{0.0016}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XLVIII@out{0.021}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XLVIII@out{0.0024}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XLVIII@out{0.017}\else\def\macros@XLVIII@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XLVIII@out}\newcommand\macros@XLIX[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@XLIX@out{\{"median": 0.27, "error plus": 0.12, "error minus": 0.1, "5th percentile": 0.17, "95th percentile": 0.39, "10th percentile": 0.18, "90th percentile": 0.37\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@XLIX@out{0.27}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@XLIX@out{0.12}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@XLIX@out{0.1}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@XLIX@out{0.17}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@XLIX@out{0.39}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@XLIX@out{0.18}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@XLIX@out{0.37}\else\def\macros@XLIX@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@XLIX@out}\newcommand\macros@L[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@L@out{\{"median": 0.053, "error plus": 0.034, "error minus": 0.04, "5th percentile": 0.013, "95th percentile": 0.085, "10th percentile": 0.023, "90th percentile": 0.075\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@L@out{0.053}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@L@out{0.034}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@L@out{0.04}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@L@out{0.013}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@L@out{0.085}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@L@out{0.023}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@L@out{0.075}\else\def\macros@L@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@L@out}\newcommand\macros@LI[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@LI@out{\{"median": 0.55, "error plus": 0.24, "error minus": 0.21, "5th percentile": 0.34, "95th percentile": 0.79, "10th percentile": 0.37, "90th percentile": 0.74\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@LI@out{0.55}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@LI@out{0.24}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@LI@out{0.21}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@LI@out{0.34}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@LI@out{0.79}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@LI@out{0.37}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@LI@out{0.74}\else\def\macros@LI@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@LI@out}\newcommand\macros@LII[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@LII@out{\{"median": 0.046, "error plus": 0.077, "error minus": 0.036, "5th percentile": 0.01, "95th percentile": 0.13, "10th percentile": 0.015, "90th percentile": 0.11\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@LII@out{0.046}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@LII@out{0.077}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@LII@out{0.036}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@LII@out{0.01}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@LII@out{0.13}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@LII@out{0.015}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@LII@out{0.11}\else\def\macros@LII@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@LII@out}\newcommand\macros@LIII[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@LIII@out{\{"median": 0.011, "error plus": 0.015, "error minus": 0.0077, "5th percentile": 0.0029, "95th percentile": 0.028, "10th percentile": 0.004, "90th percentile": 0.023\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@LIII@out{0.011}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@LIII@out{0.015}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@LIII@out{0.0077}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@LIII@out{0.0029}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@LIII@out{0.028}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@LIII@out{0.004}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@LIII@out{0.023}\else\def\macros@LIII@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@LIII@out}\newcommand\macros@LIV[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@LIV@out{\{"median": 0.31, "error plus": 0.1, "error minus": 0.11, "5th percentile": 0.2, "95th percentile": 0.41, "10th percentile": 0.22, "90th percentile": 0.39\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@LIV@out{0.31}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@LIV@out{0.1}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@LIV@out{0.11}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@LIV@out{0.2}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@LIV@out{0.41}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@LIV@out{0.22}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@LIV@out{0.39}\else\def\macros@LIV@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@LIV@out}\newcommand\macros@LV[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@LV@out{\{"median": 0.039, "error plus": 0.033, "error minus": 0.028, "5th percentile": 0.011, "95th percentile": 0.073, "10th percentile": 0.017, "90th percentile": 0.063\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@LV@out{0.039}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@LV@out{0.033}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@LV@out{0.028}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@LV@out{0.011}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@LV@out{0.073}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@LV@out{0.017}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@LV@out{0.063}\else\def\macros@LV@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@LV@out}\newcommand\macros@LVI[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@LVI@out{\{"median": 0.61, "error plus": 0.2, "error minus": 0.22, "5th percentile": 0.39, "95th percentile": 0.81, "10th percentile": 0.44, "90th percentile": 0.77\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@LVI@out{0.61}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@LVI@out{0.2}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@LVI@out{0.22}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@LVI@out{0.39}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@LVI@out{0.81}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@LVI@out{0.44}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@LVI@out{0.77}\else\def\macros@LVI@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@LVI@out}\newcommand\macros@LVII[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@LVII@out{\{"median": 0.066, "error plus": 0.096, "error minus": 0.048, "5th percentile": 0.018, "95th percentile": 0.18, "10th percentile": 0.025, "90th percentile": 0.14\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@LVII@out{0.066}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@LVII@out{0.096}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@LVII@out{0.048}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@LVII@out{0.018}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@LVII@out{0.18}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@LVII@out{0.025}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@LVII@out{0.14}\else\def\macros@LVII@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@LVII@out}\newcommand\macros@LVIII[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@LVIII@out{\{"median": 0.00024, "error plus": 0.0067, "error minus": 0.00024, "5th percentile": 1.1e{-}07, "95th percentile": 0.0083, "10th percentile": 9.2e{-}07, "90th percentile": 0.0039\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@LVIII@out{0.00024}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@LVIII@out{0.0067}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@LVIII@out{0.00024}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@LVIII@out{1.1e{-}07}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@LVIII@out{0.0083}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@LVIII@out{9.2e{-}07}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@LVIII@out{0.0039}\else\def\macros@LVIII@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@LVIII@out}\newcommand\macros@LIX[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@LIX@out{\{"median": 0.28, "error plus": 0.14, "error minus": 0.13, "5th percentile": 0.15, "95th percentile": 0.42, "10th percentile": 0.18, "90th percentile": 0.39\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@LIX@out{0.28}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@LIX@out{0.14}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@LIX@out{0.13}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@LIX@out{0.15}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@LIX@out{0.42}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@LIX@out{0.18}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@LIX@out{0.39}\else\def\macros@LIX@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@LIX@out}\newcommand\macros@LX[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@LX@out{\{"median": 0.06, "error plus": 0.031, "error minus": 0.037, "5th percentile": 0.023, "95th percentile": 0.089, "10th percentile": 0.032, "90th percentile": 0.082\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@LX@out{0.06}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@LX@out{0.031}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@LX@out{0.037}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@LX@out{0.023}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@LX@out{0.089}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@LX@out{0.032}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@LX@out{0.082}\else\def\macros@LX@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@LX@out}\newcommand\macros@LXI[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@LXI@out{\{"median": 0.55, "error plus": 0.29, "error minus": 0.26, "5th percentile": 0.3, "95th percentile": 0.85, "10th percentile": 0.35, "90th percentile": 0.78\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@LXI@out{0.55}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@LXI@out{0.29}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@LXI@out{0.26}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@LXI@out{0.3}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@LXI@out{0.85}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@LXI@out{0.35}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@LXI@out{0.78}\else\def\macros@LXI@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@LXI@out}\newcommand\macros@LXII[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@LXII@out{\{"median": 0.0015, "error plus": 0.042, "error minus": 0.0015, "5th percentile": 6.9e{-}07, "95th percentile": 0.052, "10th percentile": 5.7e{-}06, "90th percentile": 0.024\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@LXII@out{0.0015}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@LXII@out{0.042}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@LXII@out{0.0015}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@LXII@out{6.9e{-}07}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@LXII@out{0.052}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@LXII@out{5.7e{-}06}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@LXII@out{0.024}\else\def\macros@LXII@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@LXII@out}\newcommand\macros@LXIII[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@LXIII@out{\{"median": 6.2, "error plus": 0.64, "error minus": 1.1, "5th percentile": 5.1, "95th percentile": 6.8, "10th percentile": 5.4, "90th percentile": 6.7\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@LXIII@out{6.2}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@LXIII@out{0.64}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@LXIII@out{1.1}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@LXIII@out{5.1}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@LXIII@out{6.8}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@LXIII@out{5.4}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@LXIII@out{6.7}\else\def\macros@LXIII@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@LXIII@out}\newcommand\macros@LXIV[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@LXIV@out{\{"median": 15, "error plus": 5.5, "error minus": 2.7, "5th percentile": 12, "95th percentile": 21, "10th percentile": 13, "90th percentile": 19\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@LXIV@out{15}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@LXIV@out{5.5}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@LXIV@out{2.7}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@LXIV@out{12}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@LXIV@out{21}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@LXIV@out{13}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@LXIV@out{19}\else\def\macros@LXIV@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@LXIV@out}\newcommand\macros@LXV[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@LXV@out{\{"median": 45, "error plus": 9.9, "error minus": 6.2, "5th percentile": 39, "95th percentile": 55, "10th percentile": 40, "90th percentile": 52\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@LXV@out{45}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@LXV@out{9.9}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@LXV@out{6.2}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@LXV@out{39}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@LXV@out{55}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@LXV@out{40}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@LXV@out{52}\else\def\macros@LXV@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@LXV@out}\newcommand\macros@LXVI[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@LXVI@out{\{"median": 8.7, "error plus": 0.64, "error minus": 0.94, "5th percentile": 7.8, "95th percentile": 9.3, "10th percentile": 8.1, "90th percentile": 9.2\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@LXVI@out{8.7}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@LXVI@out{0.64}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@LXVI@out{0.94}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@LXVI@out{7.8}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@LXVI@out{9.3}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@LXVI@out{8.1}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@LXVI@out{9.2}\else\def\macros@LXVI@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@LXVI@out}\newcommand\macros@LXVII[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@LXVII@out{\{"median": 12, "error plus": 13, "error minus": 1.1, "5th percentile": 10, "95th percentile": 27, "10th percentile": 11, "90th percentile": 21\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@LXVII@out{12}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@LXVII@out{13}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@LXVII@out{1.1}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@LXVII@out{10}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@LXVII@out{27}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@LXVII@out{11}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@LXVII@out{21}\else\def\macros@LXVII@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@LXVII@out}\newcommand\macros@LXVIII[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@LXVIII@out{\{"median": 45, "error plus": 10, "error minus": 6.2, "5th percentile": 39, "95th percentile": 57, "10th percentile": 40, "90th percentile": 54\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@LXVIII@out{45}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@LXVIII@out{10}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@LXVIII@out{6.2}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@LXVIII@out{39}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@LXVIII@out{57}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@LXVIII@out{40}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@LXVIII@out{54}\else\def\macros@LXVIII@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@LXVIII@out}\newcommand\macros@LXIX[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@LXIX@out{\{"median": 6.2, "error plus": 0.64, "error minus": 1.1, "5th percentile": 5.1, "95th percentile": 6.8, "10th percentile": 5.4, "90th percentile": 6.7\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@LXIX@out{6.2}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@LXIX@out{0.64}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@LXIX@out{1.1}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@LXIX@out{5.1}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@LXIX@out{6.8}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@LXIX@out{5.4}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@LXIX@out{6.7}\else\def\macros@LXIX@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@LXIX@out}\newcommand\macros@LXX[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@LXX@out{\{"median": 11, "error plus": 4.6, "error minus": 0.59, "5th percentile": 11, "95th percentile": 17, "10th percentile": 11, "90th percentile": 15\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@LXX@out{11}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@LXX@out{4.6}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@LXX@out{0.59}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@LXX@out{11}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@LXX@out{17}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@LXX@out{11}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@LXX@out{15}\else\def\macros@LXX@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@LXX@out}\newcommand\macros@LXXI[1][all]{\ifnum\pdfstrcmp{#1}{all}=0\def\macros@LXXI@out{\{"median": 43, "error plus": 10, "error minus": 4.7, "5th percentile": 38, "95th percentile": 55, "10th percentile": 39, "90th percentile": 52\}}\else\ifnum\pdfstrcmp{#1}{median}=0\def\macros@LXXI@out{43}\else\ifnum\pdfstrcmp{#1}{error plus}=0\def\macros@LXXI@out{10}\else\ifnum\pdfstrcmp{#1}{error minus}=0\def\macros@LXXI@out{4.7}\else\ifnum\pdfstrcmp{#1}{5th percentile}=0\def\macros@LXXI@out{38}\else\ifnum\pdfstrcmp{#1}{95th percentile}=0\def\macros@LXXI@out{55}\else\ifnum\pdfstrcmp{#1}{10th percentile}=0\def\macros@LXXI@out{39}\else\ifnum\pdfstrcmp{#1}{90th percentile}=0\def\macros@LXXI@out{52}\else\def\macros@LXXI@out{??}\fi\fi\fi\fi\fi\fi\fi\fi\macros@LXXI@out}\makeatother